---
title: "RepData_PeerAssessment1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(ggplot2)
```

### 1. Code for reading in the dataset and/or processing the data

```{r}
data <- as.data.table(read.csv("activity.csv", header = TRUE, sep = ","))

# Add a column with the "date" column converted to a POSIXlt object
data[, pdate := as.POSIXct(date, tz = "", format = "%Y-%m-%d"), by = interval]
```

### 2. Histogram of the total number of steps taken each day

```{r}
steps_by_day <- copy(data[,c("steps", "date")])
setkey(steps_by_day, date)
steps_by_day_t <- steps_by_day[, .(total = sum(steps, na.rm = TRUE)), by = date]

hist(steps_by_day_t$total, xlab = "Steps per day", probability = FALSE, breaks = 20, main = "Histogram of Steps per Day")

```

### 3. Mean and median number of steps taken each day

Total number of steps taken per day:  
Mean  : `r mean(steps_by_day_t$total)`  
Median: `r median(steps_by_day_t$total)`

### 4. Time series plot of the average number of steps taken

```{r}
steps_by_interval <- data[, .(mean_steps = mean(steps, na.rm = TRUE)), by = interval]
plot(steps_by_interval$interval, steps_by_interval$mean_steps, type="l")
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
max_interval <- unique(steps_by_interval[mean_steps==max(steps_by_interval$mean_steps)]$interval)
```

Interval with maximum average number of steps: `r max_interval` (`r (max_interval %% 60)/5`:`r floor(max_interval / 60)`)

### 6. Code to describe and show a strategy for imputing missing data

TBD

### 7. Histogram of the total number of steps taken each day after missing values are imputed

TBD

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r chunk8}
is_weekend_day <- function(xdatetime) {
    as.factor(weekdays(xdatetime)) %in% c("Saturday", "Sunday")
}

data[, weekend := is_weekend_day(pdate)]

steps_by_interval_weekend <- data[data$weekend, .(mean_steps = mean(steps, na.rm = TRUE)), by = interval]
steps_by_interval_weekday <- data[!data$weekend, .(mean_steps = mean(steps, na.rm = TRUE)), by = interval]

library(ggpubr)
library(magrittr)

max_y_value <- max(max(steps_by_interval_weekend$mean_steps), max(steps_by_interval_weekday$mean_steps))

weekend_plot <- 
    ggplot(steps_by_interval_weekend, aes(interval, mean_steps)) + geom_line() +
    xlab("") + ylab("") + ggtitle("Weekend") +
    coord_cartesian(ylim = c(0, max_y_value)) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          plot.title = element_text(hjust = 0.5))

weekday_plot <- 
    ggplot(steps_by_interval_weekday, aes(interval, mean_steps)) + geom_line() +
    xlab("Interval") + ylab("") + ggtitle("Weekday") +
    coord_cartesian(ylim = c(0, max_y_value)) +
    theme(plot.title = element_text(hjust = 0.5))

panel_plot <- ggarrange(weekend_plot, weekday_plot, ncol = 1, nrow = 2)
panel_plot <- annotate_figure(panel_plot,
                left = text_grob("Number of Steps", color = "black", rot = 90)
)
show(panel_plot)
```

There are a few quick observations that can be made from the panel plot:

- The person is active earlier in the morning during the week
- The activity level is generally lower during the day on weekdays



